Manuscript Figures for Genome Biology

### 2018-05-25 manuscript_figures/manuscript_figures.Rmd
Add Figure 5, and a supp figure

# Figure 2D: Some methods are tailored to detecting smaller effect sizes
p4 <- covariateLinePlot(de5, alpha=0.05, covname="ind_covariate", trans="log1p",
                        nbins=25)
leg <- get_legend(p4)
p4 <- p4 + theme(legend.position="none") + 
  ggtitle("Mean % significant by covariate %ile") + 
  theme(axis.title = element_text(face="bold", size = 10),
        plot.title = element_text(face="bold", size = 14)) +
  xlab("Independent covariate percentile")

@@

This is a heatmap of methods by case studies with color representing the ranking
by of number of rejections @ FDR cutoff 0.10.

@@

Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
                            tableOnly = TRUE)$method)
Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10)
Fig4
ggsave(file.path(outdir, "Figure4a.pdf"), width=6, height=4)

Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10, 
                               tableOnly = TRUE, excludeMethods = NULL)$method)
Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
                        excludeMethods = NULL)
Fig4All
ggsave(file.path(outdir, "Figure4a_all.pdf"), width=6, height=4)

@@

```{r}
# Supplementary table S3 
library(xtable)
tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = 0.10,
                        tableOnly = TRUE)
tabS3 <- tabS3 %>% 
  mutate(mean_prop = ifelse(is.na(mean_prop), "-",
                            paste0(round(mean_prop,2), " (", round(min_prop,2), 
                            "-", round(max_prop, 2), ")"))) %>%
  select(casestudy, method, mean_prop) %>%
  spread(casestudy, mean_prop)

@@

pfdr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = 0.10, xlab = "Comparison",
                colorLow = "yellow", colorHigh = "navy",
                fill = "FDR",
                rowOrder = Fig4.rowOrder)

ptpr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = 0.10, xlab = "Comparison",
                fill = "TPR",
                rowOrder = Fig4.rowOrder)

@@

This is a multi-panel Figure displaying examples of covariate line plots in the
case studies.

```{r, Figure 5}
# TO DO
```

@@

@@

```{r}
# by covariate
wcs <- which(grepl("GWAS", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
                      function(x) x[2]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Covariate",
                excludeMethods = NULL, Nlabel = FALSE) + 
  ggtitle("GWAS")
ggsave(file.path(outdir, "FigureS1.pdf"), width=4, height=4)
```

@@

```{r}
#by dataset
wcs <- which(grepl("GSEA", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("GSEA")
ggsave(file.path(outdir, "FigureS2.pdf"), width=4, height=4)
```

@@

```{r}
wcs <- which(grepl("ChIPseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("ChIP-Seq")
ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=4)
```

### RNA-seq
```{r}
# by dataset
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("polyester|yeast"), dataset) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", 
                excludeMethods = NULL, Nlabel = FALSE) + 
  ggtitle("RNA-seq")
ggsave(file.path(outdir, "FigureS4.pdf"), width=4, height=4)
```

@@

# by dataset
pa <- plotMethodRanks(objs, colLabels = ds, alpha = 0.10, xlab = "Dataset") + 
  ggtitle("scRNA-seq")

# by method
pb <- plotMethodRanks(objs, colLabels = meth, alpha = 0.10, xlab = "Method") + 
  ggtitle("scRNA-seq")

# by covariate
pc <- plotMethodRanks(objs, colLabels = cv, alpha = 0.10, xlab = "Covariate") + 
  ggtitle("scRNA-seq")

@@

```{r}
# by dataset, excluding otu when genus is available
# excluding log-ubiquity
# excluding enigma-so4 (no discoveries)
wcs <- which(grepl("microbiome", casestudy) & 
              !(grepl(c("baxter|goodrich|papa"), dataset) & 
              grepl(c("otu"), dataset)) &
              !grepl("log-ubiquity", dataset) &
              !grepl("so4", dataset))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = 0.10, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("Microbiome")
ggsave(file.path(outdir, "FigureS6.pdf"), width=7.5, height=4)
```

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL)

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "FDR")

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "TPR")

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL)

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = 0.10, xlab = "Comparison",
                excludeMethods = NULL) 
```

>>>>

# Figure 2D: Some methods are tailored to detecting smaller effect sizes
p4 <- covariateLinePlot(de5, alpha=alpha, covname="log2FC", trans="log1p",
                        nbins=25)
leg <- get_legend(p4)
p4 <- p4 + theme(legend.position="none") + 
  ggtitle("Mean % significant by effect size %ile") + 
  theme(axis.title = element_text(face="bold", size = 10),
        plot.title = element_text(face="bold", size = 14)) +
  xlab("Effect size (log2 fold change) percentile")

@@

This is a heatmap of methods by case studies with color representing the ranking
by of number of rejections @ FDR cutoff `r alpha`.

@@

Fig4.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                            tableOnly = TRUE)$method)
Fig4 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha)
Fig4
ggsave(file.path(outdir, "Figure4a.pdf"), width=6, height=4)

Fig4All.rowOrder <- levels(plotMethodRanks(objects, colLabels = casestudy, alpha = alpha, 
                               tableOnly = TRUE, excludeMethods = NULL)$method)
Fig4All <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        excludeMethods = NULL)
Fig4All
ggsave(file.path(outdir, "Figure4a_all.pdf"), width=6, height=4)

@@

```{r}
# Supplementary table S3 
library(xtable)
tabS3 <- plotMethodRanks(objects, colLabels = casestudy, alpha = alpha,
                        tableOnly = TRUE)
tabS3 <- tabS3 %>% 
  mutate(mean_prop = ifelse(is.na(mean_prop), "-",
                            paste0(round(mean_prop,2), " (", round(min_prop,2), 
                            "-", round(max_prop, 2), ")"))) %>%
  select(casestudy, method, mean_prop) %>%
  spread(casestudy, mean_prop)

@@

pfdr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                colorLow = "yellow", colorHigh = "navy",
                fill = "FDR",
                rowOrder = Fig4.rowOrder)

ptpr <- plotMethodRanks(objs, 
                colLabels = c("sim", "sim"), 
                alpha = alpha, xlab = "Comparison",
                fill = "TPR",
                rowOrder = Fig4.rowOrder)

@@

This is a multi-panel Figure displaying examples of covariate line plots in the
case studies. We'll create a 4-panel figure with the following case studies:
RNAseq, scRNAseq, GSEA, GWAS.

```{r, Figure 5}

objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("RNAseq", objects) & grepl("brain", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pA <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("RNA-seq case study: Brain") +
  xlab("Overall mean expression percentile")

leg <-  get_legend(pA)
pA <- pA + theme(legend.position="none")

resfile <- objects[grepl("RNAseq", objects) & grepl("human", objects) &
                   grepl("mast", objects) & grepl("det", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pB <- covariateLinePlot(res, alpha=alpha, covname="detection", trans="log1p",
                        nbins=25) +
  ggtitle("scRNA-seq case study: Human, MAST") +
  xlab("Detection rate percentile") + 
  theme(legend.position="none")

resfile <- objects[grepl("GSEA", objects) & grepl("mouse", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pC <- covariateLinePlot(res, alpha=alpha, covname="ind_covariate", trans="log1p",
                        nbins=25) +
  ggtitle("GSEA case study: Mouse") +
  xlab("Gene Set size percentile") + 
  theme(legend.position="none")

resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)
pD <- covariateLinePlot(res, alpha=alpha, covname="ind_covar_AF", trans="log1p",
                        nbins=25) +
  ggtitle("GWAS case study")+
  xlab("Minor Allele Frequency (MAF) percentile") +
  theme(legend.position="none")

p5 <- plot_grid(pA, pB, pC, pD, ncol=2, labels = LETTERS[1:4])
p5 <- plot_grid(p5, leg, ncol=2, rel_widths = c(1,0.2))
ggsave(file.path(outdir, "Figure5.pdf"), width=9, height=7)

@@

## Case-study specific plot analagous to Figure 2

Figure 2 panels C-E for a case study. For now picking GWAS.

```{r Figure2.5}
objects <- list.files( path, recursive=TRUE, pattern="rds", full.names=TRUE )
resfile <- objects[grepl("GWAS", objects) & grepl("maf", objects)] 
res <- readRDS(file=resfile)
assayNames(res) <- "qvalue"
res <- addDefaultMetrics(res)

# Figure 2C: Total number of discoveries varied across method
p3 <- rejections_scatter(res, supplementary = FALSE) +
      theme(legend.position="none")+ 
  ggtitle("Number of rejections by alpha cutoff") + 
  xlab("alpha cutoff") +
  theme(axis.title = element_text(face="bold", size = 10),
        plot.title = element_text(face="bold", size = 14))

# Figure 2D: Some methods are tailored to detecting smaller effect sizes
p4 <- covariateLinePlot(res, alpha=alpha, covname="effect_size", trans="log1p",
                        transx = "exp", nbins=25)
leg <- get_legend(p4)
p4 <- p4 + theme(legend.position="none") + 
  ggtitle("Mean % significant by effect size %ile") + 
  theme(axis.title = element_text(face="bold", size = 10),
        plot.title = element_text(face="bold", size = 14)) +
  xlab("Effect size percentile")

# Figure 2E - upsetr plot - save separately
pdf(file.path(outdir, "FigureS0C.pdf"), width = 10, height = 4, onefile=FALSE)
plotFDRMethodsOverlap(res, 
                      alpha=alpha, nsets=ncol(res),
                      order.by="freq", decreasing=TRUE,
                      supplementary=FALSE, nintersects = 18) 
grid.text("Overlap for GWAS Case study, MAF covariate", x = 0.6, y = 0.97, 
          gp = gpar(fontsize = 16, fontface = "bold"))
grid.text("C", x = 0.02, y = 0.97, 
          gp = gpar(fontsize = 24, fontface = "bold")) 
dev.off()

# Figure 2 (ALL)
Fig2 <- plot_grid(p3, p4, ncol = 2, 
          labels = LETTERS[1:2], label_size = 20)
Fig2 <- plot_grid(Fig2, leg, ncol = 2, rel_widths = c(1, 0.2))
Fig2
ggsave(file.path(outdir, "FigureS0A-B.pdf"), width=10, height=4)

@@

```{r}
# by covariate
wcs <- which(grepl("GWAS", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-"), 
                      function(x) x[2]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Covariate",
                excludeMethods = NULL, Nlabel = FALSE) + 
  ggtitle("GWAS")
ggsave(file.path(outdir, "FigureS1.pdf"), width=4, height=4)
```

@@

```{r}
#by dataset
wcs <- which(grepl("GSEA", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("GSEA")
ggsave(file.path(outdir, "FigureS2.pdf"), width=4, height=4)
```

@@

```{r}
wcs <- which(grepl("ChIPseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("ChIP-Seq")
ggsave(file.path(outdir, "FigureS3.pdf"), width=5, height=4)
```

### RNA-seq
```{r}
# by dataset
wcs <- which(grepl("RNAseq", casestudy) & 
            !grepl(c("polyester|yeast"), dataset) &
            !grepl("scRNAseq", casestudy))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", 
                excludeMethods = NULL, Nlabel = FALSE) + 
  ggtitle("RNA-seq")
ggsave(file.path(outdir, "FigureS4.pdf"), width=4, height=4)
```

@@

# by dataset
pa <- plotMethodRanks(objs, colLabels = ds, alpha = alpha, xlab = "Dataset") + 
  ggtitle("scRNA-seq")

# by method
pb <- plotMethodRanks(objs, colLabels = meth, alpha = alpha, xlab = "Method") + 
  ggtitle("scRNA-seq")

# by covariate
pc <- plotMethodRanks(objs, colLabels = cv, alpha = alpha, xlab = "Covariate") + 
  ggtitle("scRNA-seq")

@@

```{r}
# by dataset, excluding otu when genus is available
# excluding log-ubiquity
# excluding enigma-so4 (no discoveries)
wcs <- which(grepl("microbiome", casestudy) & 
              !(grepl(c("baxter|goodrich|papa"), dataset) & 
              grepl(c("otu"), dataset)) &
              !grepl("log-ubiquity", dataset) &
              !grepl("so4", dataset))
objs <- objects[wcs]
labs <- unlist(lapply(str_split(dataset[wcs], pattern = "-benchmark.rds"), 
                      function(x) x[1]))
plotMethodRanks(objs, colLabels = labs, alpha = alpha, xlab = "Dataset", Nlabel = FALSE) + 
  ggtitle("Microbiome")
ggsave(file.path(outdir, "FigureS6.pdf"), width=7.5, height=4)
```

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL)

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "FDR")

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL, 
                fill = "TPR")

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL)

@@

plotMethodRanks(objs, 
                colLabels = labs, 
                alpha = alpha, xlab = "Comparison",
                excludeMethods = NULL) 
```
